WASP-4\lowercase{b} Arrived Early for the TESS Mission

### 2019-02-20 review/referee_response_0.txt

Dear Dr. Endl,

We thank you for organizing the review of our paper, and the referee
for constructive criticism.  We have considered the referee's remarks
carefully and revised our manuscript accordingly.

Below, we have reproduced the referee's report, along with our
point-by-point responses. A list of changes is also appended to the
manuscript.

Sincerely,

Luke Bouma

********************************************************************************

Referee comment:

The authors present a number of new and high-quality transit observations of
the hot Jupiter WASP-4b observed by TESS. After comparing the predicted
mid-transit times (obtained only considering the pre-existing transits in the
literature) with the measured values from the TESS light curves, the authors
conclude that the TESS transits occur ~ 82 seconds earlier than predicted. They
examine different possibilities and find that the 2 most favored scenarios to
explain this deviation are orbital decay and apsidal precession. However, the
authors caution that more observations are needed to confirm these results.
Also, they redetermine various parameters of the system, finding them in
general agreement with literature values.

TODO TODO TODO
* TODO: add more weight to the distant companion idea.
* ADD: our own analysis of the RV data, and the resulting implied period
  derivative.
* ADD: RV analysis emphasis (!!).

In general, the manuscript is scientifically accurate and the results presented
in it certainly will be of interest to other investigators in
the field. Then, I consider this paper worthy of being published after the
concerns described below be addressed.

Below are my comments, beginning with the major points. I recommend the authors
take them into consideration when revising their manuscript.

-- Major points:

1) In section 3.1, the authors claim "..We included data from peer-reviewed
literature for which the analysis was based on observations of a single,
complete transit,..". However, the transits listed in Table 2 corresponding to
the epochs E=-827 (ttra= 2454697.79817; Winn et al. 2009), E=-804 (ttra=
2454728.57767; Hoyer et al. 2013), E=-537 (ttra= 2455085.88418; Dragomir et al.
2011), E=-208 (ttra= 2455526.16356; Ranjan et al. 2014) are incomplete light
curves, and those from the epochs E=-795 (2454740.62125; Hoyer et al. 2013),
E=-251 (2455468.61943; Hoyer et al. 2013) and E=21 (2455832.61815; Hoyer et al.
2013) present scarcity of after ingress/before egress data-points. On one hand,
given that several partial light curves have been included in their study, the
use of the word "complete" to indicate the characteristics of the transits in
their sample, must be avoided. 
On the other hand, at least two other transits included in their sample show
visible anomalies, probably produced by the pass of the planet in front of one
or several cold spots during transit (E=-561, ttra=2455053.76774 and E=-526,
ttra=2455100.60595, both from Sanchis-Ojeda et al. 2011). Different studies
have proven that mid transit-times measured from partial light curves (Sada et
al. 2012, Barros et al. 2013, Nikolov et al. 2015, Mancini et al. 2018) or
light curves with anomalies due to the presence of spots (Oshagh et al. 2013,
Mazeh et al. 2015, Ioannidis et al. 2016), usually present larger uncertainties
and untrustable values. Taking in consideration the conclusions of these
previous works, it is possible that the exclusion of incomplete/with anomalies
transits might affect the outcome of their analysis and conclusions,
particularly with respect to orbital decay and apsidal precession. Then, in
order to warrant confident results, I would like to see the same analysis but
without including partial transits or with visible spot-crossing events on
them.

TODO TODO TODO
* USE: best-judgement.
* ENSURE: do not include partial transits.
* OMITTING THE ONES WITH VISIBLE SPOT-CROSSINGS: meh. Maybe focus on the ones
  with excess reisduals during partial phases.

2) As mentioned in 1), another possible interpretation for explaining the
deviation from a linear ephemeris of the TESS mid-transit times could be the
presence of anomalies in the light curves due to stellar activity. It has been
well established that WASP-4 is an active star with a rotation period (P) in
between 20-40 days (P ~ 22 days and P ~ 34 days were determined by
Sanchis-Ojeda et al. 2011 and Hoyer et al. 2013, respectively), which is
expected for a main-sequence G7 star. Given that the TESS mid-transits span
around 30 days, wouldn't it be possible that these measurements be affected by
the existence of unseen stellar spots? Regarding this point, I would like to
see an extra paragraph in the Section "Interpretation" with a discussion about
this scenario.

TODO TODO TODO
* ONE IDEA: consider ADDING the spots (e.g., from the hubble data, at the
  observed amplitude). Run it through the lightcurves, over a full transit
  phase. QUANITFY: How much does the transit mid-time change?
* ALSO QUANTIFY: what is the peak-to-peak stellar variability? (Can we confirm
  the previously measured rotation periods?)
* The period is the same order of time as the observations...

-- Minor points:

0. TITLE:

i) As it is, the title implies a very strong asseveration. It would be more
appropriate "Possible Early Arrival of WASP-4b for the TESS Mission" or
something similar.

TODO TODO TODO
* Pending spot studies: no.

1. INTRODUCTION:

i) Please, include here a reference for TESS.

  > Response: Done.

2. NEW TRANSITS AND SYSTEM PARAMETERS:

Section 2.1:

i) It could be informative to provide a reference for the "Deep Space Network".

  > Response: Barring any specific suggestion by the referee, we will leave
  > this one for the reader to pursue at their leisure.

ii) barycentric Julian date --> Barycentric Julian Date

  > Response: Done.

iii) For the casual reader, it would be important to briefly specify what
"Threshold Crossing Events" means, and what the flags "Reaction Wheel
Desaturation Event" and "Manual Exclude" represent. Footnotes including this
information would be appropriate. Also, could the authors describe what a
"Presearch Data Conditioning" lightcurve is?

  > Response: We have clarified that a "Presearch Data
  > Conditioning" lightcurve is at a stage of process in which
  > systematic trends have been removed through the decorrelation
  > process described by Smith et al (2017).
  >
  > We now omit explicit mention of "Treshold Crossing Events" in
  > favor of a more qualitative description.
  >
  > As described in the text, the flags "Reaction Wheel
  > Desaturation Event" and "Manual Exclude" together mean that the
  > data are potentially of poor quality due to "momentum dumps".
  > We referred the reader to the data release notes for an
  > explanation of what a "momentum dump" is, and have added text
  > to this footnote to improve its clarity.

iv) It might be worthwhile to mention what "ramp-like systematic effects" are
and also include a reference.

  > Response: We have changed the wording from "ramp-like
  > systematic effects" to "correlated red noise". No reference is
  > needed.

v) It would be important to include in this section an explanation of how
photometric data-point errors are estimated.

  > Response: Section 2.2, paragraph 3, final sentence explains
  > this: "The uncertainty in each data point was set equal to the
  > root-mean-square (rms) level of the out-of-transit data." We
  > have added the word "photometric" to the above sentence, to
  > clarify that we mean the photometric data points.

Section 2.2:

i) Could the authors explicitly indicate, in the first paragraph, the number of
transits observed by TESS finally used to create the phase-folded light curve?
Furthermore, the manuscript would be more readable if the main text refers
first to Figure 1 instead of Figure 2.

  > Response: We have clarified that the phase-folded lightcurve
  > comes from 18 transits.  We have left the figure ordering
  > as-is.

ii) The authors say that the phase-folded lightcurve was fitted by using the
implementation of Kreidberg (2015) of the analytic model of Mandel & Agol
(2002). Is this referring to the BATMAN code? If it is so, It would be good to
mention the code's name in the text. Also, I would appreciate if, in the main
text, the authors clarify from which source are the values used as initial
photometric parameters (planet to star radius ratio, orbital distance to
stellar radius ratio, inclination, etc). On the other hand, for highlighting
the results obtained with TESS data, it would be important to visualize all the
transit parameters derived from section 2.2 in an independent table.
Furthermore, to see the agreement between the transit parameters reported in
previous studies and those resulting from this work, it would be useful to
present together the values obtained in both cases.

  > Response: We now mention the name "BATMAN". We do the same with
  > "emcee".
  >
  > We did not explicitly describe our choice for the initial
  > parameters, because the converged best-fit values from the
  > chain do not depend on them.  For reproducibility purposes,
  > similar best-fit values can be obtained independent of initial
  > parameter choices. Barring an overruling request from the
  > referee, we will not mention our explicit choice in the text
  > (it was to use values from Hoyer et al., 2013).
  >
  > The transit parameters from the phase-folded fit: Rp/Rstar,
  > inclination, a/Rstar, and limb darkening parameters are given
  > in the first five lines of Table 1.
  >
  > The transit mid-times from the independent fits to each transit
  > are given in Table 2.
  >
  > The parameters which are not reported are the depths of the
  > individual transits, and the linear fit parameters. Since the
  > focus of this study is the transit times, we would prefer to
  > omit these additional parameters for the sake of concision.
  >
  > For the same reason, we leave the interested reader to explore
  > differences in the stellar and planetary parameters by citing
  > relevant work in the last paragraph of Section 2.3. From e.g.,
  > Table 8 of Southworth+2009, we see that our values agree with
  > those of previous studies, and so do not merit further
  > attention, given the main focus of the present work.

iii) In the caption of Figure 2, it would be important to specify which is the
bin size used to determine the yellow points. Furthermore, in Figure 2, the
photometric error of the unbinned data-points should be included.

  > Response: We have added a representative error bar to Figure 2
  > which shows the photometric error of the unbinned data points.
  > 
  > We have opted to not explicitly specify the bin size. It can be
  > inferred from inspection of the plot, and is only relevant for
  > visualization purposes.

iv) To estimate transit times for each individual lightcurve, did the authors
use the same code as for the phase-folded light curve? In this case, how were
the mid-transit times' uncertainties determined?

  > Response: This is discussed in paragraphs 3 and 4 of Section 2.2:

    > """To measure the transit times, we returned to the `cleaned' PDC
    > time series and fitted the data within four transit durations of
    > each transit separately. We used four free parameters: the time of
    > mid-transit $t_{\rm tra}$, the planet-to-star radius ratio, and the
    > slope and intercept of a linear trend to account for any slow
    > variations unrelated to the transit.  We fixed the remaining
    > parameters at the values that had been determined from the
    > phase-folded TESS lightcurve.  The uncertainty in each
    > \added{photometric} data point was set equal to the
    > root-mean-square (rms) level of the out-of-transit data.
    > 
    > To verify that the measured uncertainties are estimated accurately,
    > we computed the reduced $\chi^2$ for a linear ephemeris fit to the
    > measured TESS mid-transit times.  We found that $\chi^2 = 9.2$,
    > with $n=16$ degrees of freedom.  The variance of the $\chi^2$
    > distribution is $2n$, so we would expect $\chi^2 = 16 \pm 5.7$.
    > Visually inspecting the residuals showed that the error variance
    > had been overestimated, so we multiplied the measured TESS errors
    > by a factor $f=0.76$, forcing a reduced $\chi^2$ of unity.  This
    > lowered the mean uncertainty of the transit midtimes from $29.8$ to
    > $22.6$ seconds.  We verified that omitting this step did not
    > appreciably alter any of our conclusions."""

  > So yes, we used the same transit model as before (the BATMAN
  > code), but with the addition of the slope and intercept of a
  > linear trend as free parameters.  Our proceduring for
  > determining the mid-transit time uncertainties is also
  > described.

v) Is χ^2=9.2 the reduced chi-square? If it is so, It should be clarified by
using a different symbol or by adding a sub-index, for example. Please, fix
this in the rest of the manuscript when necessary.

  > Response: $\chi^2 = 9.2$ is the $\chi^2$ value. We have
  > corrected the typo in the text, in which we wrote "reduced
  > $\chi^2$'.

vi) χ^2 = 16 {plus minus} 5.7 --> χ^2 = 9.2 {plus minus} 5.7

  > Response: For 18 transits, and 2 degrees of freedom, we
  > would expect $\chi^2 = 16 \pm 5.7$.

vii) The note at the end of Table 2 explains what the parameter sigmat0
represents, but it's not clear which is its relation with the mid-transit
times' measurements presented in the same table.

  > Response: This was a typo -- $\sigma_{t_0}$ should have read
  > $\sigma_{t_{\rm tra}}$. We have added a sentence to summarize
  > the complete procedure in the note at the end of Table 2.

Section 2.3:

i) The authors should include the references for each of the mentioned catalogs
and also for the Yonsei Yale isochrones.

  > Response: Done, thank you.

ii) How did the authors estimate the errors of the stellar and planet
parameters?

  > Response: We have added the following sentence in the last
  > paragraph of Section 2.3: "The uncertainties in our derived
  > stellar and planetary parameters are propagated according to
  > standard analytic formulae, under the assumption of
  > uncorrelated and normally distributed initial measurements."

3. TIMING ANALYSIS

Section 3.1:

i) The sentence "..Since those data points carry significant weight in the
analysis, we checked that the timestamps in their data represent mid-exposure
times, that the barycentric correction was performed correctly, and that the
time system of the final results was BJD(TDB).." is confusing, because seems to
imply that these checks were only performed on the Huitson et al. mid-transit
times and not in all the literature mid-transit times. It would be
recommendable to clarify this point.

  > Response: we have modified the wording slightly to clarify. The point is
  > that for these Gemini South observations, we confirmed with the authors
  > through back and forth correspondence that the details of their
  > time-keeping were sound.  We made this extra effort because of the weight
  > of the Huitson+(2017) times in the analysis.
  >
  > For the remaining data points, we used the references stated in Section 3.1
  > and Table 2 to confirm that we were using with mid-exposure times, and
  > corrected from HJD_UTC to BJD_TDB as appropriate. However the point is
  > indeed that because of the weight of the Huitson et al (2017) times, we
  > made an particular extra effort to confirm their accuracy with the original
  > authors.

ii) Could the authors specify what is the meaning of each of the variables in
Equation (1)?

  > Response: Done.

Section 3.2:

i) How did you fit the mid times of the preTESS measurements? Did you use a
weighted least-square fit?

  > Response: Yes, this was through weighted least-squares. We have changed the
  > wording accordingly.

ii) Figure 3:

I would recommend changing the "title" indicated in bold font in the caption,
by a more explicit sentence about what is shown in the plot. To give an
example, a possibility could be "Observed minus predicted mid-transit times".

  > Response: Though we appreciate the suggestion, we have opted to keep the
  > "title" as is. The nuance of what "predicted" means for this plot (a linear
  > ephemeris, based only on the pre-TESS times) is sufficient that we find it
  > to be better left in the body of the caption.

Also: ii-1) In the legends of both panels, could the authors specify which is
the bin size of the TESS light curves? (Please, fix this in all the
manuscript).

  > Response: We have added the following sentence in the caption: "The binned
  > TESS point is the weighted average of 18 TESS transits".  The caption of
  > Figure 4 already stated the same.  We have changed the words "Dots" to
  > "Symbols" in the caption of Figure 6 in order to clarify this for the
  > reader.

ii-2) "The red band shows the average deviation of the TESS transits,..." -->
"The red band shows the average deviation of the TESS transits {plus
minus}1σ,..."

  > Response: Done.

iii) It would be important to introduce references for equations (4) and (5).
Also, for the casual reader, It might be informative to give a brief
explanation of what sidereal and anomalistic periods are.

  > Response: Equations (4) and (5) are generic equations for the Taylor series
  > expansion of the transit and occultation times as a function of epoch, and
  > do not require references.  We have added the following sentence to describe the
  > sidereal and anomalistic periods: "The sidereal period is the duration
  > required to return to the same orientation with respect to the stars; the
  > slightly longer anomalistic period is the duration required to reach a
  > fixed longitude with respect to the rotating line of apsides."

iv) Figure 4: In the caption, could the authors explicitly indicate what the
plot shows in each of its panels (top and bottom)?

  > Response: Done.

v) Similar to i) How did you fit the mid-transit times and occultations shown
in Figure 4?

  > Response: We have added the following sentence in Section 3.2:
  > "We fit each model by assuming a Gaussian likelihood and
  > sampling over the posterior probability distributions."

vi) In the caption of Table 4, please be more explicit, "Best-fit model
parameters" of what?

  > Response: The caption now reads "Best-fit transit timing model
  > parameters".

vii) Given that the mid-transit times values used in this work are taken from
different authors with different methodologies to perform the fitting and
detrending of the light curves, could the authors mention how this could affect
their results?

TODO TODO TODO
* We are trusting that people fit transits.
* (Maybe cite some of the Espinoza studies about detailed transit fitting
  systematics)

********************************************************************************
4. INTERPRETATION

Referee comment:

Section 4.1:

i) First line: The word "caused" is repeated twice.

  > Response: Corrected.

ii) In the caption of Figure 5 should be indicated what open circles represent.

  > Response: Done.

iii) The characteristic timescale value calculated in equation (13), is it
computed by considering a theoretical dP/dt value or from the one estimated
through the observations?

  > Response: Changed wording to remove any ambiguity.

iv) The authors mention that there are about 20 Hot Jupiters for which the
theoretical timescale is shorter, where are the timescale values obtained from?
Are they estimated from equation (13)? If this is the case, how do the
theoretical dP/dts are computed?

  > Response: As noted in the caption of Figure 5, the decay timescale values
  > are obtained from Equation 15, assuming Qstar'=10^7.  We have slightly
  > modified the text for clarity on this point.

v) In the top panels of Figures 4 and 6, the orbital decay model evolution
shows that at the beginning the O-C values increase until, at some point, they
reach a maximum and after that, they start decreasing. What is the physical
explanation for that behavior? If the orbital period is diminishing, wouldn't
it be expected to see only a decreasing in the O-C values? Why the O-C values
would increase at these first epochs?

  > Response: To make Figure 4 and 6, we start with the observed transit and
  > occultation mid-times, as a function of epoch number.  We then fit a line
  > to these data.  If the data were generated by a constant period orbit (with
  > gaussian uncertainties) the residuals to such a fit would show no
  > correlation with epoch number.  However if the data were generated by a
  > planet with a constant period derivative (with gaussian uncertainties), the
  > residuals to such a fit would always be a parabola.  This is because the
  > linear fit to the data will have contributions from the beginning, middle,
  > and end of the transit timeseries. Thus the best-fit linear ephemeris (to
  > data that actually have a decreasing period) has a best-fit period
  > somewhere between the minimum and maximum period. At the earliest epochs,
  > the best-fit period is an underestimate, so you see the increase in O-C. At
  > the latest epochs, the best-fit period is an overestimate, so you see the
  > decrease.

Section 4.2:

i) It might be informative to explicitly include the equation (14) of Ragozzine
& Wolf (2009) to see of which parameters is depending on.

  > Response: we have opted against this suggestion, chiefly to encourage the
  > interested reader to refer to the original study!

### 2019-02-20 paper/ms.tex
remove unused ETD appendix
WIP referee report update to paper
almost all of the minor changes. major suggestions still to be
incorporated, with the additional analysis of the RV data.

\documentclass[12pt,twocolumn,tighten]{aastex62}
%\pdfoutput=1 %for arXiv submission
\usepackage{amsmath,amstext,amssymb}
\usepackage[T1]{fontenc}
\usepackage{apjfonts}
\usepackage[figure,figure*]{hypcap}
\usepackage{graphics,graphicx}
\usepackage{hyperref}
\usepackage{comment}

@@

%\NewPageAfterKeywords

@@

% \author[0000-0003-1963-9616]{D. A. Caldwell} % douglas.caldwell@nasa.gov % WAITING

@@

% \author{M. Fausnaugh} % faus@mit.edu WAITING
% \affiliation{ Department of Physics and Kavli Institute for Astrophysics
% and Space Research, Massachusetts Institute of Technology, Cambridge, MA
% 02139, USA }

@@

% \author{N. Guerrero} % nmg@mit.edu WAITING
% \affiliation{ Department of Physics and Kavli Institute for Astrophysics
% and Space Research, Massachusetts Institute of Technology, Cambridge, MA
% 02139, USA }

@@

% \author{B. Wohler} % bill.wohler@nasa.gov WAITING

\begin{abstract}
    The Transiting Exoplanet Survey Satellite (TESS) recently observed
    18 transits of the hot Jupiter WASP-4b.  The sequence of transits
    occurred $81.6 \pm 11.7$~seconds earlier than had been predicted,
    based on data stretching back to 2007.  This is unlikely to be the
    result of a clock error, because TESS observations of other hot
    Jupiters (WASP-6b, 18b, and 46b) are compatible with a constant
    period, ruling out an 81.6-second offset at the 6.4$\sigma$ level.
    The 1.3-day orbital period of WASP-4b appears to be decreasing at
    a rate of $\dot{P} = -12.6 \pm 1.2$ milliseconds per year.  The
    apparent period change might be caused by tidal orbital decay or
    apsidal precession, although both interpretations have
    shortcomings.  The gravitational influence of a third body is
    another possibility, although at present there is no evidence for
    such a body. Further observations are needed to confirm and
    understand the timing variation.
\end{abstract}

@@

Although the Transiting Exoplanet Survey Satellite (TESS) is designed
to detect new planets, it is also performing precise photometric
monitoring of most of the planets that have been discovered by
ground-based transit surveys over the last two decades.  One of the
applications of the new TESS data is to search for timing anomalies in
hot Jupiter systems.  Long-term monitoring of hot Jupiter transit and
occultation times should eventually reveal two distinct processes.

@@

\begin{figure*}[t]
    \begin{center}
        \includegraphics[width=0.9\textwidth]{f2.png}
    \end{center}
    \vspace{-0.7cm}
    \caption{
        {\bf Phase-folded lightcurve of WASP-4b.} Black points are
        TESS data. Yellow points are binned measurements.  The bottom
        panel shows the residuals.  The fit to the phase-folded
        transit (blue line) is used when measuring mid-transit times
        for the individual transits (see text).
        \label{fig:phasefold}
    }
\end{figure*}

@@

WASP-4 was observed by TESS with Camera 2 from August 23 to September
20, 2018, within the second ``sector'' of science operations.  The
star is designated as TIC 402026209 in the TESS Input Catalog
\citep{stassun_TIC_2018}.  The pixel data for an $11\times11$ array
surrounding WASP-4 were averaged into 2-minute stacks by the onboard
computer.  The data were downlinked via the Deep Space Network, and
the spacecraft timestamps were calibrated against the ground-station
clocks.  The spacecraft clock times were then transformed by the
Payload Operations Center into the {\it Temps Dynamique Barycentrique}
(TDB) reference system.  The images were then reduced to lightcurves
by the Science Processing Operations Center (SPOC) at NASA
Ames~\citep{jenkins_tess_2016}.  During this processing, the SPOC used
the known spacecraft trajectory to compute the barycentric time
corrections on a target by target basis, and expressed the timestamps
as Barycentric TESS Julian Dates (BTJD), which is simply the
barycentric Julian date minus 2{,}457{,}000.  Lightcurves that were
flagged as Threshold Crossing Events by the SPOC pipeline were then
vetted and released by the MIT TESS Science Office to the Mikulski
Archive for Space Telescopes on November 29,
2018~\citep{ricker_tess_alerts_2018}.

We began our analysis with the Presearch Data Conditioning (PDC)
lightcurve \citep{smith_kepler_apertures_2017,smith_kepler_PDC_2017}.
We processed the lightcurve as follows.  First, we removed all points
with non-zero quality flags.  This removed data that might have been
adversely affected by ``momentum dumps,'' the firing of thrusters and
resetting of reaction
wheels\footnote{\url{archive.stsci.edu/missions/tess/doc/tess_drn/tess_sector_02_drn02_v01.pdf}}
that took place every 2.5 days during sector 2
\citep{tess_data_product_description_2018}.  The data during these
events were assigned quality flags corresponding to ``Reaction Wheel
Desaturation Event'' and ``Manual Exclude''.  For WASP-4, these flags
were simultaneously set for 54 distinct cadences, and there were 10
momentum dumps, averaging about 10 minutes of flagged data per dump.
Out of caution, we clipped out an additional 10 minutes before and
after every momentum dump.  We also removed the data within the first
and last hours of both orbits, because of ramp-like systematic effects
that appear during those time ranges.

@@

Using the cleaned PDC lightcurve, we used the Box Least Squares
algorithm \citep{kovacs_box-fitting_2002} to estimate the orbital
period, transit duration, and a reference epoch using the TESS data
alone.  Based on the results, we isolated the data within 4 transit
durations of each transit midpoint.  To find the transit parameters
that best fit the data, we first fitted a line to the out-of-transit
flux measurements surrounding each transit, and divided it out.  We
then created a phase-folded lightcurve and fitted a standard transit
model using the analytic formulae given by
\citet{mandel_analytic_2002} and implemented by
\citet{kreidberg_batman_2015}.  We assumed the orbit to be circular,
consistent with the limits from radial velocities and occultation
timing
\citep{beerer_secondary_2011,knutson_friends_2014,bonomo_gaps_2017}.
The free parameters were the reference epoch, the planet to star
radius ratio $R_{\rm p}/R_\star$, the orbital distance to stellar
radius ratio $a/R_\star$, the inclination $i$, two quadratic
limb-darkening coefficients $(u_{\rm linear}, u_{\rm quad})$, and the
orbital period $P$.

We sampled the posterior probability distribution for all the
parameters using the algorithm proposed by
\citet{goodman_ensemble_2010} and implemented by
\citet{foreman-mackey_emcee_2013}.  Table~1 gives the results, which
are in reasonable agreement with the parameters reported by {\it
e.g.}, \citet{southworth_high-precision_2009} and
\citet{huitson_gemini_2017}.  Figure~\ref{fig:phasefold} shows the
phase-folded lightcurve.

To measure the transit times, we returned to the `cleaned' PDC time
series and fitted the data within four transit durations of each
transit separately. We used four free parameters: the time of
mid-transit $t_{\rm tra}$, the planet-to-star radius ratio, and the
slope and intercept of a linear trend to account for any slow
variations unrelated to the transit.  We fixed the remaining
parameters at the values that had been determined from the
phase-folded TESS lightcurve.  The uncertainty in each data point was
set equal to the root-mean-square (rms) level of the out-of-transit
data.

To verify that the measured uncertainties are estimated accurately, we
computed the reduced $\chi^2$ for a linear ephemeris fit to the
measured TESS mid-transit times.  We found that $\chi^2 = 9.2$, with
$n=16$ degrees of freedom.  The variance of the $\chi^2$ distribution
is $2n$, so we would expect $\chi^2 = 16 \pm 5.7$.  Visually
inspecting the residuals showed that the error variance had been
overestimated, so we multiplied the measured TESS errors by a factor
$f=0.76$, forcing a reduced $\chi^2$ of unity.  This lowered the mean
uncertainty of the transit midtimes from $29.8$ to $22.6$ seconds.  We
verified that omitting this step did not appreciably alter any of our
conclusions.

@@

If the timing variation is being caused caused entirely by orbital
decay, then the characteristic timescale of the decay is
\begin{equation}
  \frac{P}{ \dot{P} } = 9.2 \, {\rm Myr}.
\end{equation}
For comparison, the corresponding time for WASP-12b is 3.2\,Myr
\citep{patra_2017}.

>>>>

\documentclass[12pt,twocolumn,tighten,trackchanges]{aastex62}
%\pdfoutput=1 %for arXiv submission
\usepackage{amsmath,amstext,amssymb}
\usepackage[T1]{fontenc}
\usepackage{apjfonts}
\usepackage[figure,figure*]{hypcap}
\usepackage{graphics,graphicx}
\usepackage{hyperref}
\usepackage{comment}

@@

%\NewPageAfterKeywords
\listofchanges

@@

% \author[0000-0003-1963-9616]{D. A. Caldwell} % douglas.caldwell@nasa.gov % 
%%%WAITING (LIKELY YES)

@@

\added{\author{M. Fausnaugh} % faus@mit.edu CONFIRMED
02139, USA }}

@@

\added{
\author{N. Guerrero} % nmg@mit.edu CONFIRMED
}

@@

\added{
\author{B. Wohler} % bill.wohler@nasa.gov YES
}

\begin{abstract}
    The Transiting Exoplanet Survey Satellite (TESS) recently observed
    18 transits of the hot Jupiter WASP-4b.  The sequence of transits
    occurred $81.6 \pm 11.7$~seconds earlier than had been predicted,
    based on data stretching back to 2007.  This is unlikely to be the
    result of a clock error, because TESS observations of other hot
    Jupiters (WASP-6b, 18b, and 46b) are compatible with a constant
    period, ruling out an 81.6-second offset at the 6.4$\sigma$ level.
    The 1.3-day orbital period of WASP-4b appears to be decreasing at
    a rate of $\dot{P} = -12.6 \pm 1.2$ milliseconds per year.  The
    apparent period change might be caused by tidal orbital decay or
    apsidal precession, although both interpretations have
    shortcomings.  The gravitational influence of a third body is
    another possibility, although at present there is \replaced{no}{minimal} 
    evidence for
    such a body. Further observations are needed to confirm and
    understand the timing variation.
\end{abstract}

@@

Although the Transiting Exoplanet Survey Satellite (TESS, \added{
\citealt{ricker_transiting_2015}}) is designed
to detect new planets, it is also performing precise photometric
monitoring of most of the planets that have been discovered by
ground-based transit surveys over the last two decades.  One of the
applications of the new TESS data is to search for timing anomalies in
hot Jupiter systems.  Long-term monitoring of hot Jupiter transit and
occultation times should eventually reveal two distinct processes.

@@

\begin{figure*}[t]
    \begin{center}
        \includegraphics[width=0.9\textwidth]{f2.png}
    \end{center}
    \vspace{-0.7cm}
    \caption{
        {\bf Phase-folded lightcurve of WASP-4b.} Gray points are
        TESS \replaced{data}{flux measurements, with median 
        $1\sigma$ error shown in the lower 
        right.} 
        Yellow points are binned measurements.  The bottom
        panel shows the residuals.  The fit to the phase-folded
        transit (blue line) is used when measuring mid-transit times
        \replaced{for the}{of} individual transits (see 
        \S~\ref{sec:measurement}).
        \label{fig:phasefold}
    }
\end{figure*}

@@

WASP-4 was observed by TESS with Camera 2 from August 23 to September
20, 2018, within the second ``sector'' of science operations.  The
star is designated as TIC 402026209 in the TESS Input Catalog
\citep{stassun_TIC_2018}.  The pixel data for an $11\times11$ array
surrounding WASP-4 were averaged into 2-minute stacks by the onboard
computer.  The data were downlinked via the Deep Space Network, and
the spacecraft timestamps were calibrated against the ground-station
clocks.  The spacecraft clock times were then transformed by the
Payload Operations Center into the {\it Temps Dynamique Barycentrique}
(TDB) reference system.  The images were then reduced to lightcurves
by the Science Processing Operations Center (SPOC) at NASA
Ames~\citep{jenkins_tess_2016}.  During this processing, the SPOC used
the known spacecraft trajectory to compute the barycentric time
corrections on a target by target basis, and expressed the timestamps
as Barycentric TESS Julian Dates (BTJD), which is simply the
\replaced{b}{B}arycentric Julian \replaced{d}{D}ate minus 2{,}457{,}000.  
Lightcurves that 
were
flagged \deleted{as Threshold Crossing Events} by the SPOC pipeline \added{as 
crossing a transit detection threshold} were then
vetted and released by the MIT TESS Science Office to the Mikulski
Archive for Space Telescopes on November 29,
2018~\citep{ricker_tess_alerts_2018}.

We began our analysis with the Presearch Data Conditioning (PDC)
lightcurve\replaced{ 
\citep{smith_kepler_apertures_2017,smith_kepler_PDC_2017}}{, which has had 
systematic trends removed 
through the decorrelation process described by 
\citet{smith_kepler_apertures_2017,smith_kepler_PDC_2017}}.
We \added{then} processed the lightcurve as follows.  First, we removed all 
points
with non-zero quality flags.  This removed data that might have been
adversely affected by ``momentum dumps,'' the firing of thrusters and
resetting of reaction
wheels\footnote{\added{Details concerning the spacecraft pointing and momentum 
dumps are described in the data release 
notes:}\url{archive.stsci.edu/missions/tess/doc/tess_drn/tess_sector_02_drn02_v01.pdf}}
that took place every 2.5 days during sector 2
\deleted{\citep{tess_data_product_description_2018}}.  The data during these
events were assigned quality flags corresponding to ``Reaction Wheel
Desaturation Event'' and ``Manual Exclude'' 
\added{\citep[][Table 28]{tess_data_product_description_2018}}.  For WASP-4, 
these flags
were simultaneously set for 54 distinct cadences, and there were 10
momentum dumps, averaging about 10 minutes of flagged data per dump.
Out of caution, we clipped out an additional 10 minutes before and
after every momentum dump.  We also removed the data within the first
and last hours of both orbits, because of \replaced{ramp-like systematic 
effects that appear}{correlated red noise that appears} during those time 
ranges.

@@

Using the cleaned PDC lightcurve, we applied the Box Least Squares
algorithm \citep{kovacs_box-fitting_2002} to estimate the orbital
period, transit duration, and a reference epoch using the TESS data
alone.  Based on the results, we isolated the data within 4 transit
durations of each transit midpoint.  To find the transit parameters
that best fit the data, we first fitted a line to the out-of-transit
flux measurements surrounding each transit, and divided it out.  We
then created a phase-folded lightcurve \added{from all 18 transits}
and fitted a standard transit
model using the analytic formulae given by
\citet{mandel_analytic_2002} and implemented by
\replaced{\citet{kreidberg_batman_2015}}{\citet[][\texttt{BATMAN}]{kreidberg_batman_2015}}
We assumed 
the orbit to be circular,
consistent with the limits from radial velocities and occultation
timing
\citep{beerer_secondary_2011,knutson_friends_2014,bonomo_gaps_2017}.
The free parameters were the reference epoch, the planet to star
radius ratio $R_{\rm p}/R_\star$, the orbital distance to stellar
radius ratio $a/R_\star$, the inclination $i$, two quadratic
limb-darkening coefficients $(u_{\rm linear}, u_{\rm quad})$, and the
orbital period $P$.

We sampled the posterior probability distribution for all the
parameters using the algorithm proposed by
\citet{goodman_ensemble_2010} and implemented by
\replaced{\citet{foreman-mackey_emcee_2013}}{\citet[][\texttt{emcee}]{foreman-mackey_emcee_2013}}.
  Table~1 gives the 
results, which
are in reasonable agreement with the parameters reported by {\it
e.g.}, \citet{southworth_high-precision_2009} and
\citet{huitson_gemini_2017}.  Figure~\ref{fig:phasefold} shows the
phase-folded lightcurve.

To measure the transit times, we returned to the `cleaned' PDC time
series and fitted the data within four transit durations of each
transit separately. We used four free parameters: the time of
mid-transit $t_{\rm tra}$, the planet-to-star radius ratio, and the
slope and intercept of a linear trend to account for any slow
variations unrelated to the transit.  We fixed the remaining
parameters at the values that had been determined from the
phase-folded TESS lightcurve.  The uncertainty in each \added{photometric} 
data point was set equal to the root-mean-square (rms) level of the 
out-of-transit data.

To verify that the measured uncertainties are estimated accurately, we
computed the \deleted{reduced} $\chi^2$ \added{value} for a linear ephemeris 
fit to the
measured TESS mid-transit times.  We found that $\chi^2 = 9.2$, with
$n=16$ degrees of freedom.  The variance of the $\chi^2$ distribution
is $2n$, so we would expect $\chi^2 = 16 \pm 5.7$.  Visually
inspecting the residuals showed that the error variance had been
overestimated, so we multiplied the measured TESS errors by a factor
$f=0.76$, forcing a reduced $\chi^2$ of unity.  This lowered the mean
uncertainty of the transit midtimes from $29.8$ to $22.6$ seconds.  We
verified that omitting this step did not appreciably alter any of our
conclusions.

@@

If the timing variation is \deleted{being caused} caused entirely by
orbital decay, then \replaced{the}{the best-fit model parameters yield a}
characteristic \replaced{timescale of the decay is}{decay timescale of}
\begin{equation}
  \frac{P}{ \dot{P} } = 9.2 \, {\rm Myr}.
\end{equation}
For comparison, the corresponding time for WASP-12b is 3.2\,Myr
\citep{patra_2017}.
